---
title: "Assignment 2b"
format: docx
editor: visual
knitr: 
  opts_chunk: 
    echo: true
    results: 'hide'
---

## **STEP 1: Setting my working directory and loading required packages in R**

```{r}
rm(list = ls())         #clears environment 
directory <- setwd("~/Documents/SFU courses/Spring 2022/Econ836 - A. Econometrics/Assignm") #sets working directory 

library(haven)          #needed to load the .dta stata data file
library(tidyverse)      #needed for datacleaning 
library(dplyr)
library(quantreg)       #for running regressions 
library(labelled)       #needed for labels
library(modelsummary)   #needed for tables
library(data.table) 
library(stargazer)
```

**STEP 2: Loading the dataset**

Here I will load the *2016 Census Public Use Microdata File (PUMF) needed for the project*

```{r}
d.s <- read_dta("~/Documents/SFU courses/Spring 2022/Econ836 - A. Econometrics/Assignm/Census_2016_Individual_PUMF.dta")
```

**STEP 3: Cleaning the dataset**

Here I will clean all 8 variables that are needed in my analysis which have missing data and require a reference level for the dummies:

1.  EmpIn: this is the variable that defines the "income" and has data that is not available is labelled as "8,888,888" and data that is not applicable is labelled as "99,999,999" so here I will replace those observations with NA

2.  agegrp: this is the variable that defines the "agegroup" and has data that is not available is labelled as "88", so here I will replace those observations with NA.

    The "left out" category for agegroup is "25-29 years" which is coded as 9 in the data, so I put this as my reference level

3.  MatStH: this is the variable that defines "martial status" and There are no unavailable data in this category.

    The "left out" category for "martial status" is the never-married individuals which is coded as 1 in the data, so I put this as my reference level

4.  hdegree: this is the variable that defines "education" and and has data that is not available is labelled as "88" and is not applicable is labelled as "99", so here I will replace those observations with NA.

    The "left out" category for "education" are individuals with a highschool degree which is coded as 2 in the data, so I put this as my reference level

5.  hhsize:this is the variable that defines "household size" and and has data that is not available is labelled as "8", so here I will replace those observations with NA.

6.  kol: this is the variable that defines "knowledge language" and this has data that is not available labelled as "8", so here I replace those observations with NA.

    The "left out" category for "knowledge language" are individuals with "English Only" which is coded as 1 in the data, so I put this as my reference level

7.  cma: this defines "*census metropolitan areas"* and here the left out catagory are the"non-cmas" which is coded as 999, so I put this as the reference level

8.  abor: I create this variable to correctly capture the aboriginal people

```{r}
#filtering out "Income" (income)-----------------------------------------------------------------
class(d.s$EmpIn) #checking the class of the variable: numeric
d.s <- d.s %>% rename(income=EmpIn) %>% 
  mutate(income = case_when( 
    income %in% c(88888888,99999999) ~ as.numeric(NA),
    TRUE ~ as.numeric(income)
  ))

#filtering out "Agegroup" (agegrp) ----------------------------------------------------------------
class(d.s$agegrp) #checking the class of the variable: numeric 
d.s <- d.s %>%
  mutate(agegrp = case_when(
    agegrp == 88 ~ as.double(NA),
    TRUE ~ as.double(agegrp)
  ))
d.s$agegrp <- factor(d.s$agegrp) #converting agegroup to factor/catagorical variable 

d.s$agegrp <- relevel(d.s$agegrp, ref=9) #taking the correct reference level (25-29)
d.s <- d.s %>% rename(age=agegrp) #renaming agegrp to age 


#converting MarStH to a factor and releveling it (marital)--------------------------------------
class(d.s$MarStH)
d.s$MarStH <- factor(d.s$MarStH) 
d.s$MarStH <- relevel(d.s$MarStH, ref=1) #taking the correct reference level (single people)
d.s <- d.s %>% rename(marital=MarStH)      #renaming MarStH to marital


#filtering out education (school)--------------------------------------------------------------

class(d.s$hdgree)
d.s <- d.s %>% 
  mutate(hdgree = case_when( 
    hdgree %in% c(88,99) ~ as.double(NA),
    TRUE ~ as.double(hdgree)
  ))
d.s$hdgree <- factor(d.s$hdgree)
d.s$hdgree <- relevel(d.s$hdgree, ref=2) #taking the correct reference level (highschool) 
d.s <- d.s %>% rename(school=hdgree)      #renaming hdgree to school 

#filtering out household size (hhsize)------------------------------------------------------------

class(d.s$hhsize) 

d.s <- d.s %>% 
  mutate(hhsize = case_when( 
    hhsize ==8 ~ as.double(NA),
    TRUE ~ as.double(hhsize)
  ))
d.s$hhsize <- factor(d.s$hhsize)
levels(d.s$hhsize) #checking the level

#filtering the language catagory (lang) -------------------------------------------------------

class(d.s$kol) #checking the class: double
d.s <- d.s %>% 
  mutate(kol = case_when( 
    kol ==8 ~ as.double(NA),
    TRUE ~ as.double(kol)
  ))


d.s$kol <- factor(d.s$kol)
d.s$kol <- relevel(d.s$kol, ref=1) #setting the correct reference level (English only)
d.s <- d.s %>% rename(lang=kol)  #renaming kol to lang

#releveling CMAs -------------------------------------------------------------------

class(d.s$cma)

d.s$cma <- factor(d.s$cma)
d.s$cma <- relevel(d.s$cma, ref='999')


#making the abor variable to include all catagories of indians ------------------------------

#ethder=1 are the 'North American Aboriginal origins'
#regind=1 are the 'registered Indians' 
d.s <- 
  d.s %>%
  filter(ethder != 88) %>% 
  mutate(
    abor = case_when(
      ethder != 1 ~ ethder,
      ethder == 1 & regind == 1 ~ as.double(999), # 999 is Registered Indians,
      ethder == 1 & regind == 0 & aboid != 6 ~ aboid,
      ethder == 1 & regind == 0 & aboid == 6 ~ as.double(998) # 998 is Ancestry
    )
  )

# Now give them labels 

d.s <-
  d.s %>% 
  mutate(abor = 
    case_when(
      abor == 999 ~ 'Registered Indian',
      abor == 100 ~ 'North American Indian',
      abor == 200 ~ 'Métis',
      abor == 300  ~ 'Inuit',
      abor == 400  ~ 'Multiple',
      abor == 500  ~ 'Other aboriginal response',
      abor == 600 ~ 'Ancestry',
      TRUE ~ as.character(ethder),
    ),
    abor = as.factor(abor) %>% relevel('4')
  )

#Get the accurate sample size by filtering out only individuals who work for someone else 

d.s <-
  d.s %>% 
  filter(income>0, cow==1)

```

**Step 4: Get the correct samples for the regressions**

```{r}
#WHOLE OF CANADA

men_can <-
  d.s %>% 
  filter(Sex == 2)

women_can <- 
  d.s %>% 
  filter(Sex ==1)

# MONTREAL

men_montreal <-
  d.s %>% 
  filter(cma == 462,
         Sex == 2)

women_montreal <-
  d.s %>% 
  filter(cma == 462,
         Sex == 1)

# TORONTO

men_toronto <-
  d.s %>% 
  filter(cma == 535,
         Sex == 2)

women_toronto <-
  d.s %>% 
  filter(cma == 535,
         Sex == 1)

# WINNIPEG

men_winnipeg <-
  d.s %>% 
  filter(cma == 535,
         Sex == 2)

women_winnipeg <-
  d.s %>% 
  filter(cma == 602,
         Sex == 1)

#CALGARY

men_calgary <-
  d.s %>% 
  filter(cma == 825,
         Sex == 2)

women_calgary <-
  d.s %>% 
  filter(cma == 825,
         Sex == 1)

# EDMONTON

men_edmonton <-
  d.s %>% 
  filter(cma == 835,
         Sex == 2)

women_edmonton <-
  d.s %>% 
  filter(cma == 835,
         Sex == 1)

# VANCOUVER

men_vancouver <-
  d.s %>% 
  filter(cma == 933,
         Sex == 1)

women_vancouver <-
  d.s %>% 
  filter(cma == 933,
         Sex == 2)
```

**Step 5: Run the regressionsRRr5**

```{r}
#Canada 

reg_canada_men <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_can,
     tau = c(0.2,0.5,0.8,0.9))

summary(reg_canada_men, se="nid") #se="nid" gives us t-stat and p value
reg_canada_women <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_can,
     tau = c(0.2,0.5,0.8,0.9))

#Men pseudo R-squared

fit0 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_can)
fit1 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_can)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R1 <- 1 - fit1$rho/fit0$rho


#Women pseudo R-squared
fit2 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_can)
fit3 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_can)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R2 <- 1 - fit3$rho/fit2$rho

#Montreal

reg_men_montreal <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_montreal,
     tau = c(0.2,0.5,0.8,0.9))

reg_women_montreal <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_montreal,
     tau = c(0.2,0.5,0.8,0.9))


#men pseudo R-squared
fit4 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_montreal)
fit5 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_montreal)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R3 <- 1 - fit5$rho/fit4$rho

#women pseudo R-squared 

fit6 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_montreal)
fit7 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_montreal)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R4 <- 1 - fit7$rho/fit6$rho

#Toronto 

reg_men_toronto <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_toronto,
     tau = c(0.2,0.5,0.8,0.9))

reg_women_toronto <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_toronto,
     tau = c(0.2,0.5,0.8,0.9))

#men pseudo R-squared
fit8 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_toronto)
fit9 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_toronto)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R5 <- 1 - fit9$rho/fit8$rho

#women pseudo R-squared 

fit10 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_toronto)
fit11 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_toronto)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R6 <- 1 - fit7$rho/fit6$rho

#Winnipeg

reg_men_winnipeg <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_winnipeg,
     tau = c(0.2,0.5,0.8,0.9))

reg_women_winnipeg <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_winnipeg,
     tau = c(0.2,0.5,0.8,0.9))

#men pseudo R-squared
fit12 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_winnipeg)
fit13 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_winnipeg)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R7 <- 1 - fit13$rho/fit12$rho

#women pseudo R-squared 

fit14 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_winnipeg)
fit15 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_winnipeg)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R8 <- 1 - fit15$rho/fit14$rho
#Calgary 

reg_men_calgary <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_calgary,
     tau = c(0.2,0.5,0.8,0.9))

reg_women_calgary <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_calgary,
     tau = c(0.2,0.5,0.8,0.9))

#men pseudo R-squared
fit16 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_calgary)
fit17 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_calgary)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R9 <- 1 - fit17$rho/fit16$rho

#women pseudo R-squared 

fit18 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_calgary)
fit19 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_calgary)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R10 <- 1 - fit19$rho/fit18$rho

#Edmonton 

reg_men_edmonton <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_edmonton,
     tau = c(0.2,0.5,0.8,0.9))

reg_women_edmonton <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_edmonton,
     tau = c(0.2,0.5,0.8,0.9))

#men pseudo R-squared
fit20 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_edmonton)
fit21 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_edmonton)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R11 <- 1 - fit21$rho/fit20$rho

#women pseudo R-squared 

fit22 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_edmonton)
fit23 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_edmonton)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R12 <- 1 - fit23$rho/fit22$rho

#Vancouver 

reg_men_vancouver <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = men_vancouver,
     tau = c(0.2,0.5,0.8,0.9))

reg_women_vancouver <-
  rq(log(income) ~ abor + marital + age + hhsize + school + lang, 
     data = women_vancouver,
     tau = c(0.2,0.5,0.8,0.9))

#men pseudo R-squared
fit24 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = men_vancouver)
fit25 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = men_vancouver)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R13 <- 1 - fit25$rho/fit24$rho

#women pseudo R-squared 

fit26 <- rq(log(income)~1,tau=c(0.2,0.5,0.8,0.9),data = women_vancouver)
fit27 <- rq(log(income)~abor,tau=c(0.2,0.5,0.8,0.9),data = women_vancouver)

rho <- function(u,tau=.5)u*(tau - (u < 0))
R14 <- 1 - fit27$rho/fit26$rho
```

**Step 5: Creating output to extract relevant coefficients for each table**

```{r}
#when youre taking the coefs it is already a data.frame 

#canada 

write.csv(reg_canada_men$coef, "1.reg_canada_men.csv")

write.csv(reg_canada_women$coef, "2.reg_canada_women.csv")

#montreal 

write.csv(reg_men_montreal$coef, "3.reg_men_montreal.csv") 

write.csv(reg_women_montreal$coef, "4.reg_women_montreal.csv") 

#toronto 

write.csv(reg_men_toronto$coef, "5.reg_men_toronto.csv") 

write.csv(reg_women_toronto$coef, "6.reg_women_toronto.csv") 

#winnipeg reg_men_winnipeg

write.csv(reg_men_winnipeg$coef, "7.reg_men_winnipeg.csv") 

write.csv(reg_women_winnipeg$coef, "8.reg_women_winnipeg.csv") 

#calgary 

write.csv(reg_men_calgary$coef, "9.reg_men_calgary.csv") 
write.csv(reg_women_calgary$coef, "10.reg_women_calgary.csv") 

#edmonton reg_men_edmonton

write.csv(reg_men_edmonton$coef, "11.reg_men_edmonton.csv") 
write.csv(reg_women_edmonton$coef, "12.reg_women_edmonton.csv") 

#vancouver reg_men_vancouver


write.csv(reg_men_vancouver$coef, "13.reg_men_vancouver.csv") 
write.csv(reg_women_vancouver$coef, "14.reg_women_vancouver.csv") 
```
